The Way Forward in Parallelizing Dynamic Languages

### 2014-05-02 talk/icooolps2014/position-paper.tex
add citation

\subsection{Why is there a GIL?}
The GIL is a very simple synchronisation mechanism for supporting
multithreading in the interpreter. The basic guarantee is that the
GIL may only be released in-between bytecode instructions. The
interpreter can thus rely on complete isolation and atomicity of these
instructions. Additionally, it provides the application with a
sequential consistency model. As a consequence, applications can rely
on certain operations to be atomic and that they will always be
executed in the order in which they appear in the code. While
depending on this may not always be a good idea, it is done in
practice. A GIL-replacement should therefore uphold these
guarantees, while preferably also be as easily implementable as a GIL
for the interpreter.
[xxx mention that the interpreter is typically very large and maintained
by open-source communities]

@@

\begin{table*}[!ht]
  \centering
  \begin{tabular}{|l|c|c|c|c|c|}
    \hline
    & \textbf{GIL} & \textbf{Fine-grained locking}
    & \textbf{Shared-nothing} & \textbf{HTM} & \textbf{STM}\\
    \hline
    Performance (single threaded) & ++   & +  & ++   & ++ & -{-} \\
    \hline
    Performance (multithreaded)   & -{-} & +  & +    & +  & +    \\
    \hline
    Existing applications         & ++   & ++ & -{-} & ++ & ++   \\
    \hline
    Better synchronisation        & -    & -  & -    & -  & ++   \\
    \hline
    Implementation                & ++   & -  & ++   & ++ & ++   \\
    \hline
    External libraries            & ++   & ++ & ++   & ++ & ++   \\
    \hline
  \end{tabular}
  \caption{Comparison between the approaches (-{-}/-/o/+/++)}
  \label{tab:comparison}
\end{table*}

@@

\acks

Acknowledgements...

@@

>>>>

\subsection{Why is there a GIL?}
The GIL is a very simple synchronisation mechanism for supporting
multithreading in the interpreter. The basic guarantee is that the GIL
may only be released in-between bytecode instructions. The interpreter
can thus rely on complete isolation and atomicity of these
instructions. Additionally, it provides the application with a
sequential consistency model\cite{lamport79}. As a consequence,
applications can rely on certain operations to be atomic and that they
will always be executed in the order in which they appear in the
code. While depending on this may not always be a good idea, it is
done in practice. A GIL-replacement should therefore uphold these
guarantees, while preferably also be as easily implementable as a GIL
for the interpreter.
[xxx mention that the interpreter is typically
  very large and maintained by open-source communities]

@@

\begin{table*}[h]
  \centering
  \begin{tabular}{|l|c|c|c|c|c|}
    \hline
    & \textbf{GIL} & \textbf{Fine-grained locking}
    & \textbf{Shared-nothing} & \textbf{HTM} & \textbf{STM}\\
    \hline
    Performance (single threaded) & ++   & +  & ++   & ++ & -{-} \\
    \hline
    Performance (multithreaded)   & -{-} & +  & +    & +  & +    \\
    \hline
    Existing applications         & ++   & ++ & -{-} & ++ & ++   \\
    \hline
    Better synchronisation        & -    & -  & -    & -  & ++   \\
    \hline
    Implementation                & ++   & -  & ++   & ++ & ++   \\
    \hline
    External libraries            & ++   & ++ & ++   & ++ & ++   \\
    \hline
  \end{tabular}
  \caption{Comparison between the approaches (-{-}/-/o/+/++)}
  \label{tab:comparison}
\end{table*}

@@

%% \acks

%% Acknowledgements...

@@

\bibitem{lamport79}
  Lamport, Leslie. "How to make a multiprocessor computer that
  correctly executes multiprocess programs." \emph{Computers, IEEE
    Transactions} on 100.9 (1979): 690-691.
